---
title: "Aides Drug"
author: "Jonathan Urbano"
date: "June 14, 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r}
Charlie <- filter(triUmp, DrugUser == "yes")
```

```{r eval=FALSE}
Charlie <- Charlie %>%
mutate(across(where(is.character), as.factor))
```

# Introduction
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

I need to use a filter command form the helper cheat sheet. The filter to get the data out of triUmp that is comprise of drug users. I will proceed to do this in the next set of codes.


```{r}
DRUGUSR<-filter(triUmp,DrugUser== "yes")
```

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

#### Plot barplots ggplot way

```{r}
dd2 <- DRUGUSR %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

 The first plot shows how many people got infected. It looks like it is maybe less than 100 people with AIDS. In the %plot it treats both groups in % terms. 25% of people that got AIDS have taken the drugs. 75% of people that got AIDS have got the placebo. It presents that shows the drug is working from getting AIDS. The relationship is 3rd quaters of the people have got the placebo the ration was 50/50 so the drug is preventing those who took it from getting AIDS.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=DRUGUSR)
rowPerc(table1)
colPerc(table1)
```

Of the people who got AIDS 24.29% have taken the drug 75.71% have taken the placebo. From the people that haven't got anything was 47.59% to 52.41%. This is a sign of a drug that looks like it is working

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The probability of getting this result is an extremely low number of .000000423. This extremely small numer tells me that we reject the hypothesis and tells us that the drug works. Small p value shows something is happening.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
